---
title: "HW 4"
author: "Team 2"
date: "April 12, 2019"
output:
  html_document:
    pdf_document: default
    theme: cosmo
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r, echo = FALSE}
#install.packages(c("broom", "ggplot2", "dplyr", "tidyr", "corrplot", "randomForest", "olsrr", "caret", "fastDummies", "car", "pROC", "pscl"))

require(broom)
require(ggplot2)
require(dplyr)
require(tidyr)
require(corrplot)
require(randomForest)
require(olsrr)
require(caret)
require(fastDummies)
require(car)
require(pROC)
require(pscl)
# library(dummies)
# head(dummy.data.frame(data = as.data.frame(HairEyeColor),sep="."),3)
```



```{r, echo = FALSE}
training <- as.data.frame(read.csv("https://raw.githubusercontent.com/simplymathematics/621/master/HW4/insurance_training_data.csv"))
test <- as.data.frame(read.csv("https://raw.githubusercontent.com/simplymathematics/621/master/HW4/insurance_training_data.csv"))
str(training)
summary(training)
sapply(training, function(y) sum(length(which(is.na(y)))))/nrow(training)*100
# head(training)
```


Data Cleaning 

1. Removing $ sign and commas

2. Cleaning data set from  "_z" characters

3. Removing rows with NAs

```{r, echo = FALSE}
# removing "_z" character
training <- data.frame(lapply(training, function(x){
  gsub("z_", "", x)
  
}))

# removing "<" character
training <- data.frame(lapply(training, function(x){
  gsub("<", "", x)
  
}))
```

```{r, echo = FALSE}
# Removing $ sign and commas
training$HOME_VAL<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$HOME_VAL)))))
training$INCOME<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$INCOME)))))
training$BLUEBOOK<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$BLUEBOOK)))))
training$OLDCLAIM<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$OLDCLAIM)))))
```

```{r, echo = FALSE}
# removing rows with NAs
training<-na.omit(training)

summary(training)
sapply(training, function(y) sum(length(which(is.na(y)))))/nrow(training)*100
```


Data Transformation 

1. Transforming the following variables to factors

```{r, echo = FALSE}
training$TARGET_FLAG<-as.factor(training$TARGET_FLAG)
training$KIDSDRIV<-as.factor(training$KIDSDRIV)
training$HOMEKIDS<-as.factor(training$HOMEKIDS)

training_BINS<-training
```


# Model Building

##MODEl 1

1. Building logistic regression model using variables without any transformations.

```{r, echo = FALSE}

model_1<-step(glm(TARGET_FLAG~., training %>% select(-INDEX, -TARGET_AMT), family = "binomial"), direction = "backward")
summary(model_1)
vif(model_1)
```

Checking model_1 for the presence of influential values.

```{r echo=F}
plot(model_1, which = 4, id.n = 5)
model.data <- augment(model_1) %>%
  mutate(index = 1:n())
model.data %>% top_n(5, .cooksd)
ggplot(model.data, aes(index, .std.resid)) +
  geom_point(aes(color = TARGET_FLAG), alpha = .5) +
  theme_bw()
model.data %>%
  filter(abs(.std.resid) > 3)
```

Eliminating the row from training data set with influential value.

```{r fig.height=10}
training_no_infl <-training %>%
  filter(!(INDEX==4958 & AGE==44))
```

##MODEl 2

2. Building a model with removed influential values.

```{r, echo = FALSE}
model_2<-step(glm(TARGET_FLAG~., training_no_infl %>% select(-INDEX, -TARGET_AMT), family = "binomial"), direction = "backward")
summary(model_2)
vif(model_2)
```

##MODEl 3

3. Building a model with removed JOB variable to avoid multicollinearity.

```{r, echo = FALSE}
model_3<-step(glm(TARGET_FLAG~., training %>% select(-INDEX, -TARGET_AMT, -JOB), family = "binomial"), direction = "backward")
summary(model_3)
vif(model_3)
```


##MODEl 4

4. Taking log10 of INCOME and HOME_VAL to help model get the pattern better in one unit increas of of INCOME or HOME_VAL

```{r, echo = FALSE}
training$INCOME<- log10(training$INCOME+1)
training$HOME_VAL<- log10(training$HOME_VAL+1)

model_4<-step(glm(TARGET_FLAG~., training %>% select(-INDEX, -TARGET_AMT), family = "binomial"), direction = "backward")
summary(model_4)
vif(model_4)
```

##MODEl 5

5. Using derived variable - wealth accumulation which is INCOME/AGE and removing INCOME to avoid multicollinearity. 

```{r, echo = FALSE}
training$INCOME<- as.integer(training$INCOME)
training$AGE<- as.integer(training$AGE)
training$wealth_acc<-training$INCOME/training$AGE
model_5<-step(glm(TARGET_FLAG~., training %>% select(-INDEX, -TARGET_AMT, -INCOME), family = "binomial"), direction = "backward")
summary(model_5)
vif(model_5)
```

##MODEl 6

6. Creating bins for the continious variables: AGE, YOJ, INCOME, TRAVTIME, BLUEBOOK, TIF, OLDCLAIM, CAR_AGE, HOME_VAL and converting them to factors.

```{r, echo = FALSE}
# bins for HOME_VAL
training_BINS$HOME_VAL<- as.integer(training_BINS$HOME_VAL)
training_BINS$HOME_VAL_BINS<-ifelse(training_BINS$HOME_VAL < 1000, "<1000", ifelse(training_BINS$HOME_VAL < 2000, "1000-2000", ifelse(training_BINS$HOME_VAL< 3000,"2000-3000", ifelse(training_BINS$HOME_VAL< 4000,"3000-4000",ifelse(training_BINS$HOME_VAL< 5000,"4000-5000","5000+")))))

training_BINS$HOME_VAL_BINS<-as.factor(training_BINS$HOME_VAL_BINS)

# bins for AGE
training_BINS$AGE<-as.integer(training_BINS$AGE)
training_BINS$AGE_BINS<-ifelse(training_BINS$AGE < 20, "<20", ifelse(training_BINS$AGE < 30, "20-30", ifelse(training_BINS$AGE< 40,"30-40", ifelse(training_BINS$AGE< 50,"40-50",ifelse(training_BINS$AGE< 60,"50-60",ifelse(training_BINS$AGE< 70,"60-70","70+"))))))

training_BINS$AGE_BINS<-as.factor(training_BINS$AGE_BINS)

# bins for INCOME
training_BINS$INCOME<-as.integer(training_BINS$INCOME)

training_BINS$INCOME_BINS<-ifelse(training_BINS$INCOME < 1000, "<1000",ifelse(training_BINS$INCOME < 20000, "1000-20000", ifelse(training_BINS$INCOME < 40000, "20000-40000", ifelse(training_BINS$INCOME< 60000,"40000-60000", ifelse(training_BINS$INCOME< 80000,"60000-80000",ifelse(training_BINS$INCOME< 100000,"80000-100000",ifelse(training_BINS$INCOME< 120000,"100000-120000","120000+")))))))

training_BINS$INCOME_BINS<-as.factor(training_BINS$INCOME_BINS)

# bins for YOJ
training_BINS$YOJ<- as.integer(training_BINS$YOJ)

training_BINS$YOJ_BINS<-ifelse(training_BINS$YOJ < 5, "<5",ifelse(training_BINS$YOJ < 10, "5-10", ifelse(training_BINS$YOJ < 15, "10-15", ifelse(training_BINS$YOJ< 20,"15-20","20+"))))

training_BINS$YOJ_BINS<-as.factor(training_BINS$YOJ_BINS)

# bins for TRAVTIME
training_BINS$TRAVTIME<- as.integer(training_BINS$TRAVTIME)

training_BINS$TRAVTIME_BINS<-ifelse(training_BINS$TRAVTIME < 5, "<5",ifelse(training_BINS$TRAVTIME < 25, "5-25", ifelse(training_BINS$TRAVTIME < 45, "25-45", ifelse(training_BINS$TRAVTIME< 65,"45-65", ifelse(training_BINS$TRAVTIME< 85,"65-85",ifelse(training_BINS$TRAVTIME< 105,"85-105",ifelse(training_BINS$TRAVTIME< 125,"105-125",ifelse(training_BINS$TRAVTIME< 145,"125-145","145+"))))))))

training_BINS$TRAVTIME_BINS<-as.factor(training_BINS$TRAVTIME_BINS)

# bins for BLUEBOOK
training_BINS$BLUEBOOK<- as.integer(training_BINS$BLUEBOOK)

training_BINS$BLUEBOOK_BINS<-ifelse(training_BINS$BLUEBOOK < 1500, "<1500",ifelse(training_BINS$BLUEBOOK < 11500, "1500-11500", ifelse(training_BINS$BLUEBOOK < 21500, "11500-21500", ifelse(training_BINS$BLUEBOOK< 31500,"21500-31500", ifelse(training_BINS$BLUEBOOK< 41500,"31500-41500",ifelse(training_BINS$BLUEBOOK< 51500,"41500-51500",ifelse(training_BINS$BLUEBOOK< 61500,"51500-61500",ifelse(training_BINS$BLUEBOOK< 71500,"61500-71500","71500+"))))))))

training_BINS$BLUEBOOK_BINS<-as.factor(training_BINS$BLUEBOOK_BINS)

# bins for TIF
training_BINS$TIF<- as.integer(training_BINS$TIF)

training_BINS$TIF_BINS<-ifelse(training_BINS$TIF < 1, "<1",ifelse(training_BINS$TIF < 5, "1-5", ifelse(training_BINS$TIF < 10, "5-10", ifelse(training_BINS$TIF< 15,"10-15", ifelse(training_BINS$TIF< 20,"15-20",ifelse(training_BINS$TIF< 25,"20-25","25+"))))))

training_BINS$TIF_BINS<-as.factor(training_BINS$TIF_BINS)

# bins for OLDCLAIM
training_BINS$OLDCLAIM<- as.integer(training_BINS$OLDCLAIM)

training_BINS$OLDCLAIM_BINS<-ifelse(training_BINS$OLDCLAIM < 500, "<500",ifelse(training_BINS$OLDCLAIM < 1000, "500-1000", ifelse(training_BINS$OLDCLAIM < 1500, "1000-1500", ifelse(training_BINS$OLDCLAIM< 2000,"1500-2000", ifelse(training_BINS$OLDCLAIM< 2500,"2000-2500",ifelse(training_BINS$OLDCLAIM< 3000,"2500-3000","3000+"))))))

training_BINS$OLDCLAIM_BINS<-as.factor(training_BINS$OLDCLAIM_BINS)

# bins for CAR_AGE

training_BINS$CAR_AGE<- as.integer(training_BINS$CAR_AGE)

training_BINS$CAR_AGE_BINS<-ifelse(training_BINS$CAR_AGE < 1, "<1",ifelse(training_BINS$CAR_AGE < 5, "1-5", ifelse(training_BINS$CAR_AGE < 10, "5-10", ifelse(training_BINS$CAR_AGE< 15,"10-15", ifelse(training_BINS$CAR_AGE< 20,"15-20",ifelse(training_BINS$CAR_AGE< 25,"20-25","25+"))))))

training_BINS$CAR_AGE_BINS<-as.factor(training_BINS$CAR_AGE_BINS)

str(training_BINS)
```


```{r, echo = FALSE}
model_6<-step(glm(TARGET_FLAG~., training_BINS %>% select(-INDEX,-TARGET_AMT, -AGE, -YOJ, -TRAVTIME, -BLUEBOOK,-OLDCLAIM,-CAR_AGE, -TIF, -HOME_VAL, -INCOME), family = "binomial"), direction = "backward")

summary(model_6)
vif(model_6)
```

#Assessing selected models

Cleaning Test Data (test data set should have same variables, order of variables, format of variables and type of variables as a training set for model testing)

```{r, echo = FALSE}
# removing "_z" character
test <- data.frame(lapply(test, function(x){
  gsub("z_", "", x)
  
}))

# removing "<" character
test <- data.frame(lapply(test, function(x){
  gsub("<", "", x)
  
}))
```

```{r, echo = FALSE}
# Removing $ sign and commas
test$HOME_VAL<-as.numeric(gsub(",","",(gsub("\\$","",as.character(test$HOME_VAL)))))
test$INCOME<-as.numeric(gsub(",","",(gsub("\\$","",as.character(test$INCOME)))))
test$BLUEBOOK<-as.numeric(gsub(",","",(gsub("\\$","",as.character(test$BLUEBOOK)))))
test$OLDCLAIM<-as.numeric(gsub(",","",(gsub("\\$","",as.character(test$OLDCLAIM)))))
```

```{r, echo = FALSE}
# removing rows with NAs
test<-na.omit(test)
sapply(test, function(y) sum(length(which(is.na(y)))))/nrow(training)*100
```

## AIC, BIC, LOglik, pseudoR2

```{r echo=F}
m1<-cbind(AIC=AIC(model_1),BIC=BIC(model_1), loglik=logLik(model_1),pseudoR2 = pR2(model_1)[4])
m2<-cbind(AIC=AIC(model_2),BIC=BIC(model_2), loglik=logLik(model_2),pseudoR2 = pR2(model_2)[4])
m3<-cbind(AIC=AIC(model_3),BIC=BIC(model_3), loglik=logLik(model_3),pseudoR2 = pR2(model_3)[4])
m4<-cbind(AIC=AIC(model_4),BIC=BIC(model_4), loglik=logLik(model_4),pseudoR2 = pR2(model_4)[4])
m5<-cbind(AIC=AIC(model_5),BIC=BIC(model_5), loglik=logLik(model_5),pseudoR2 = pR2(model_5)[4])
m6<-cbind(AIC=AIC(model_6),BIC=BIC(model_6), loglik=logLik(model_6),pseudoR2 = pR2(model_6)[4])
summary = rbind(m1, m2, m3, m4, m5, m6)
rownames(summary) <- c("model_1","model_2","model_3","model_4","model_5","model_6")
summary
```

## Choosing the best model and applying it on test data set

```{r, echo = FALSE}
test$TRAVTIME<- as.integer(test$TRAVTIME)
# test$TIF<-as.integer(test$TIF)

test$predict_prob<-predict(model_1,test, type='response')
test$predict_target<-ifelse(test$predict_prob>=0.5,1,0)
head(test)
```

## Assessing the performance of the choosen model: ROC and AUC

```{r, echo = FALSE}
test$predict<-predict(model_1, test, type='response')
par(pty = "s")
pROC <- roc(test$TARGET_FLAG, test$predict,
smooth=FALSE, plot = TRUE, print.auc=TRUE,legacy.axes =TRUE,  col="red", main ="ROC Curve")

save(list = ls(all= TRUE), file = "logistic_object.Rdata")
```


