---
title: "HW 4"
author: "Team 2"
date: "April 12, 2019"
output:
  html_document:
    pdf_document: default
    theme: cosmo
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Overview

In this homework assignment, we will explore, analyze and model a data set containing approximately 8000
records representing a customer at an auto insurance company. Each record has two response variables. The
first response variable, TARGET_FLAG, is a 1 or a 0. A "1" means that the person was in a car crash. A zero
means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero
if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero

List of variables:

|VARIABLE_NAME  |DEFINITION                                         |THEORETICAL_EFFECT                                                                   |
|---------------|---------------------------------------------------|-------------------------------------------------------------------------------------|
|INDEX         |Identification Variable (do not use)                |None                                                                                 |
|TARGET_FLAG   |Was Car in a crash? 1=YES 0=NO                      |None                                                                                 |
|TARGET_AMT    |If car was in a crash, what was the cost            |None                                                                                 |
|AGE           |Age of Driver                                       |Very young people tend to be risky. Maybe very old people also.                      |
|BLUEBOOK      |Value of Vehicle                                    |Unknown effect on probability of collision, but probably effect the payout if there i| |              |                                                    | crash                                                                               |
|CAR_AGE       |Vehicle Age                                         |Unknown effect on probability of collision, but probably effect the payout if there i| |              |                                                    |  crash                                                                              |
|CAR_TYPE      |Type of Car                                         |Unknown effect on probability of collision, but probably effect the payout if there i| |              |                                                    |crash                                                                                |
|CAR_USE       |Vehicle Use                                         |Commercial vehicles are driven more, so might increase probability of collision      |
|CLM_FREQ      |# Claims (Past 5 Years)                             |The more claims you filed in the past, the more you are likely to file in the future |
|EDUCATION     |Max Education Level                                 |Unknown effect, but in theory more educated people tend to drive more safely         |
|HOMEKIDS      |# Children at Home                                  |Unknown effect                                                                       |
|HOME_VAL      |Home Value                                          |In theory, home owners tend to drive more responsibly                                |
|INCOME        |Income                                              |In theory, rich people tend to get into fewer crashes                                |
|JOB           |Job Category                                        |In theory, white collar jobs tend to be safer                                        |
|KIDSDRIV      |# Driving Children                                  |When teenagers drive your car, you are more likely to get into crashes               |
|MSTATUS       |Marital Status                                      |In theory, married people drive more safely                                          |
|MVR_PTS       |Motor Vehicle Record Points                         |If you get lots of traffic tickets, you tend to get into more crashes                |
|OLDCLAIM      |Total Claims (Past 5 Years)                         |If your total payout over the past five years was high, this suggests future payouts | |              |                                                    | will be high                                                                        |
|PARENT1       |Single Parent                                       |Unknown effect                                                                       |
|RED_CAR       |A Red Car                                           |Urban legend says that red cars (especially red sports cars) are more risky. Is that  | |              |                                                    |   true?                                                                              |
|REVOKED       |License Revoked (Past 7 Years)                      |If your license was revoked in the past 7 years, you probably are a more risky driver.|
|SEX           |Gender                                              |Urban legend says that women have less crashes then men. Is that true?                |
|TIF           |Time in Force                                       |People who have been customers for a long time are usually more safe.                 |
|TRAVTIME      |Distance to Work                                    |Long drives to work usually suggest greater risk                                      |
|URBANICITY    |Home/Work Area                                      |Unknown                                                                               |
|YOJ           |Years on Job                                        |People who stay at a job for a long time are usually more safe                        |

## Objective 

Our objective is to build multiple linear regression and binary logistic regression models on the training data
to predict the probability that a person will crash their car and also the amount of money it will cost if the person
does crash their car.

## Dependencies

Replication of our work requires the following packages in Rstudio:

```{r, echo = TRUE, message=FALSE, warning=FALSE, error=FALSE, comment=FALSE}
#install.packages('corrplot')
#install.packages('randomForest')  

require(ggplot2)
require(dplyr)
require(tidyr)
require(corrplot)
require(randomForest)
require(olsrr)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, comment=FALSE}
# Requirements for formatting and augmenting default settings for chunks. 
require(knitr)
require(kableExtra)
require(default)

knitr::opts_chunk$set(echo = T, message = T, warning = T) # change message/warning to F when finished

default(kable_styling)  <- list(bootstrap_options = c("basic"), 
                                position = "center", 
                                full_width = TRUE,
                                font_size = NULL)
```

# Data Exploration

First, we read the data as a csv then performed some simple statistical calculations so that we could explore the data. Below we can see a sample of the data output as it was read from the csv.

```{r, echo = FALSE}
training <- as.data.frame(read.csv("insurance_training_data.csv"))
test <- as.data.frame(read.csv("insurance-evaluation-data.csv"))

training %>% slice(1:5) %>% kable() %>% kable_styling()
```


## Summary Statistics 

We look at summary of the data

```{r, echo = FALSE}

training<-training[,-1]

summary(training)%>% kable() %>% kable_styling()

```

1) We can see that some variables have missing values.
2) Home values and Income with blank values, we imputed with mean - maybe should be 0 instead
2) We will impute with mean. 
3) We will also make variable in dollar format numeric.
4) One record has -3 for Car Age
5) There are 4 records where Job is blank

```{r, echo = FALSE}

training$TARGET_FLAG<-as.factor(training$TARGET_FLAG)

training$HOME_VAL<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$HOME_VAL)))))

training$INCOME<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$INCOME)))))

training$BLUEBOOK<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$BLUEBOOK)))))

training$OLDCLAIM<-as.numeric(gsub(",","",(gsub("\\$","",as.character(training$OLDCLAIM)))))

for(i in c(4,6,7,9,24)){
  training[is.na(training[,i]), i] <- mean(training[,i], na.rm = TRUE)
}

training$CAR_AGE[training$CAR_AGE<0]

training$JOB[training$JOB=='']
```

## Density

```{r,echo = FALSE, fig.width=8, fig.height=6}

training%>%
  select_if(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density() 
```

```{r,echo = FALSE, fig.width=5, fig.height=3}

ggplot(training, aes(x=TARGET_FLAG, y=log10(TARGET_AMT+1))) + 
  geom_violin()


ggplot(training, aes(x=MSTATUS, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=SEX, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=EDUCATION, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=CAR_USE, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=CAR_TYPE, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=RED_CAR, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=REVOKED, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=URBANICITY, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=JOB, y=log10(TARGET_AMT+1))) + 
  geom_violin()

ggplot(training, aes(x=PARENT1, y=log10(TARGET_AMT+1))) + 
  geom_violin()
```

## Scatter plot matrix

We then build scatter plot matrix for continious variables

```{r,, echo = FALSE, fig.width=12, fig.height=8,warning=FALSE,message=FALSE}

#install.packages('psych')

library(psych)


training%>%select_if(is.numeric)%>%pairs.panels( 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE  # show density plots
             )

```

## Correlation  

We can see our correlation matrix below. A dark blue circle represents a strong positive relationship and a dark red circle represents a strong negative relationship between two variables. 
```{r, echo = FALSE}
results1 <- training%>%
 select_if(is.numeric) %>% cor(method = 'pearson', use = 'complete.obs')
corrplot::corrplot(results1, method = 'circle')


```

Finally, we can use the `randomforest` package to verify our assumptions from the correlation plot.

```{r, echo = FALSE,fig.width=8, fig.height=6}
training2 <- training
training2$TARGET_AMT<- NULL
training2$TARGET_FLAG<- NULL
target1 <- training$TARGET_AMT
target2 <- training$TARGET_FLAG

fit1 <- randomForest(training2, target1, importance = TRUE, ntree = 50)

varImpPlot(fit1)

fit2 <- randomForest(training2, target2, importance = TRUE, ntree = 50)

varImpPlot(fit2)

```


# Data Preparation
In the following section we will prepare and transform our variables for our model

Replacing null value by the means

```{r}
training$AGE[is.na(training$AGE)] <- mean(training$AGE, na.rm=TRUE)
training$YOJ[is.na(training$YOJ)] <- mean(training$YOJ, na.rm=TRUE)
training$HOME_VAL[is.na(training$HOME_VAL)] <- mean(training$HOME_VAL, na.rm=TRUE)
training$CAR_AGE[is.na(training$CAR_AGE)] <- mean(training$CAR_AGE, na.rm=TRUE)
training$INCOME[is.na(training$INCOME)] <- mean(training$INCOME, na.rm=TRUE)
```

Get the complete cases
```{r}
training<- training[complete.cases(training),]
train2<-training
head(train2)
```

Based on the data preparation we transform data using log for skewed HOMEKIDS, MVR_PTS, OLDCLAIM, TIF, KIDSDRIVE, CLM_FREQ
```{r}
training$HOMEKIDS <- log(training$HOMEKIDS+1)
training$MVR_PTS <- log(training$MVR_PTS+1)
training$OLDCLAIM <- log(training$OLDCLAIM+1)
training$TIF <- log(training$TIF+1)
training$KIDSDRIV <- log(training$KIDSDRIV+1)
training$CLM_FREQ <- log(training$CLM_FREQ+1)
training <- training[, !(colnames(training) %in% c("INDEX"))]
head(training)
```


Let's select all the numeric value and study their correlation
```{r}
training_num<-select_if(training, is.numeric)
corrplot(cor(training_num), method="circle")
```

```{r}
cor.test(training_num$HOMEKIDS,training_num$AGE,method="pearson")
```

Our transformed data looks like this




