---
title: "Data_621_HW1"
author: "Olga Shiligin"
date: "14/02/2019"
output: html_document
---

Introduction
-------------


Loading necessary package

```{r, echo = FALSE}
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(car, quietly = TRUE)
library(corrplot, quietly = TRUE)
library(Hmisc, quietly = TRUE)
library(psych, quietly = TRUE)
library (MASS, quietly = TRUE)
library(lmtest, quietly = TRUE)
library(faraway, quietly = TRUE)
```

1-2. Data Exploration and Data Preparation

a. Mean/St dev/ Median

```{r, echo = FALSE}
train_data <- "moneyball-training-data.csv"
test_data <- "moneyball-evaluation-data.csv"
moneyball_data <- read.csv(train_data, header=TRUE, stringsAsFactors=FALSE, fileEncoding="latin1")
test_data <- read.csv(test_data, header = TRUE, stringsAsFactors = FALSE)
```

b. missing variables and missing values handling process

```{r}
sapply(moneyball_data, function(y) sum(length(which(is.na(y)))))/nrow(moneyball_data)*100
sapply(test_data, function(y) sum(length(which(is.na(y)))))/nrow(moneyball_data)*100
```

removing "index" and "TEAM_BATTING_HBP" columns as "TEAM_BATTING_HBP" has 92% of missing values"

```{r}
moneyball_data<-subset(moneyball_data, select = -c(INDEX))
moneyball<-subset(moneyball_data, select = -c(TEAM_BATTING_HBP))

test_data <- subset(test_data, select = -c(INDEX))
test_data <- subset(test_data, select = -c(TEAM_BATTING_HBP))
```

replacing other missing values with mean

```{r}
replace_mean <- function(x){
  x <- as.numeric(as.character(x))
  x[is.na(x)] = mean(x, na.rm=TRUE)
  return(x)
}

moneyball_filled <- apply(moneyball, 2, replace_mean)
moneyball_filled <- as.data.frame(moneyball_filled)

test_filled <- apply(test_data, 2, replace_mean)
test_filled <- as.data.frame(test_data)


```


c. Bar Chart/ Box Plot/ Histogram


```{r, echo = FALSE}
par(mfrow = c(3,5))
plot(density(moneyball_filled$TARGET_WINS))
plot(density(moneyball_filled$TEAM_BATTING_H))
plot(density(moneyball_filled$TEAM_BATTING_2B))
plot(density(moneyball_filled$TEAM_BATTING_3B))
plot(density(moneyball_filled$TEAM_BATTING_HR))
plot(density(moneyball_filled$TEAM_BATTING_BB))
plot(density(moneyball_filled$TEAM_BATTING_SO))
plot(density(moneyball_filled$TEAM_BASERUN_SB))
plot(density(moneyball_filled$TEAM_BASERUN_CS))
plot(density(moneyball_filled$TEAM_PITCHING_H))
plot(density(moneyball_filled$TEAM_PITCHING_HR))
plot(density(moneyball_filled$TEAM_PITCHING_BB))
plot(density(moneyball_filled$TEAM_PITCHING_SO))
plot(density(moneyball_filled$TEAM_FIELDING_E))
plot(density(moneyball_filled$TEAM_FIELDING_DP))
```

Target variable plot

```{r}
plot(moneyball_filled$TARGET_WINS)
hist(moneyball_filled$TARGET_WINS)
boxplot(moneyball_filled$TARGET_WINS)
summary(moneyball_filled$TARGET_WINS)
```

d. correlation variables to each other and to target

Checking for variable dependencies, as all variables are numeric we will rely on correleation.

```{r}
corr_moneyball<- cor(moneyball_filled)
round(corr_moneyball, digits =3)
corrplot(corr_moneyball, method = "circle")
```

- poor correlation with target (p<0.1): TEAM_FIELDING_E, TEAM_BASERUN_CS, TEAM_BATTING_SO, TEAM_BATTING_3B
   
- strong correlation between each others (>0.6): TEAM_PITCHING_HR vs TEAM_BATTING_HR (0.969); TEAM_BATTING_HR VS TEAM_BATTING_SO (0.693), TEAM_BATTING_3B VS TEAM_BATTING_SO (-0.656)

Possible variables to remove: TEAM_FIELDING_E, TEAM_BASERUN_CS, TEAM_BATTING_SO, TEAM_BATTING_3B, TEAM_BATTING_HR


3. Modeling
------------

Backward elimination method: 

```{r}
model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
```

Removing TEAM_PITCHING_BB

```{r}
model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
```

Removing TEAM_BASERUN_CS

```{r}
model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB  + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
```

Removing TEAM_PITCHING_HR

```{r}
model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB  + TEAM_PITCHING_H + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
```


Checking for non-linearity
--------------------------

```{r}
crPlots(model)
```

TEAM_PITCHING_H, TEAM_PITCHING_SO did not pass the check for non-linerity.


Let's transform TEAM_PITCHING_H, TEAM_PITCHING_SO and re-fit the model


```{r}
moneyball_filled$TEAM_PITCHING_H<- log10(moneyball_filled$TEAM_PITCHING_H+0.1)
moneyball_filled$TEAM_PITCHING_SO<- log10(moneyball_filled$TEAM_PITCHING_SO+0.1)

model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB  + TEAM_PITCHING_H + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
crPlots(model)
```
Removing TEAM_BATTING_SO with p-value > 0.05

```{r}
model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BASERUN_SB  + TEAM_PITCHING_H + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
```

Eliminating extreme values
--------------------------
(using Cook's distance plot as 4/(n-k-1), identify D values > cutoff)

```{r}
cutoff<-4/((nrow(moneyball_filled)-length(model$coefficients)-2))
plot(model, which = 4, cook.levels = cutoff)
plot(model, which = 5, cook.levels = cutoff)
moneyball_filled<-moneyball_filled[-which(rownames(moneyball_filled)
                                          %in% c ("1828","1342","2233")),]
```


refit the model:

```{r}
model<- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB  + TEAM_PITCHING_H + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_filled)
summary(model)
```

Check and eliminate further extrimes if any

```{r}
cutoff<-4/((nrow(moneyball_filled)-length(model$coefficients)-2))
plot(model, which = 4, cook.levels = cutoff)
plot(model, which = 5, cook.levels = cutoff)

```



Checking for multicolliniarity
------

```{r}
vif(model)
plot(model)
```

TEAM_FIELDING_E is withing the range 5-10, but eliminating TEAM_FIELDING_E does not improve the model.

```{r}
model_basic<-lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB  + TEAM_PITCHING_H + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP, moneyball_data)
summary(model_basic)
```

## Model Evaluation


